### 2018-05-08 peer-reveiw/review_one.tex
created peer review folder | started review_one

\documentclass{article}
    \usepackage[utf8]{inputenc}

    \title{\Large{Peer Review} \\
    \huge\textbf{{A Modern Look at Introductory Programming Languages}}}
    \author{Brigel Pineti}
    \date{May 2018}

    \begin{document}

    \maketitle

    \section*{}
    \textbf{1.} In order to keep the reader in track with what is happening throughout the paper, the writers may clearly indicate in the beginning of each section, the purpose of every following subsection.

    \end{document}