### 2018-05-08 peer-reveiw/review_one.tex
finished first question

\documentclass{article}
    \usepackage[utf8]{inputenc}

    \title{\Large{Peer Review} \\
    \huge\textbf{{A Modern Look at Introductory Programming Languages}}}
    \author{Brigel Pineti}
    \date{May 2018}

    \begin{document}

    \maketitle

    \section*{}
    \textbf{1.} In order to keep the reader in track with what is happening throughout the paper, the writers may clearly indicate in the beginning of each section, the purpose of every subsection. The current version of the paper contains only one section (\textit{Criteria}) besides the usual \textit{Introduction} section. The reader is being implicitly informed one the content of the \textit{Criteria} section. Sentence after sentence the content of the section is being revealed. However, it is still a bit vague and I would suggest the writers on each sentence to explicitly state the subsection number and a brief description on the intent of that subsection. I believe, the same approach should be applied to each section, including the \textit{Introduction}. Nevertheless, this cannot be achieved yet since there are supplementary sections to be added.

    \end{document}